Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-06-25 files/Chapter1/1DModel.tex
Overview v1
Veranderinge 10-06

\subsection{Equations of motion and constitutive equations}\label{ssec:1D_Model:EquationsOfMotion+ConstitutiveEquation}
In this section we introduce the Timoshenko beam theory for the transverse vibration of a uniform beam. For a reference of the model, \cite{Inm94} and \cite{Fu65} were used.\\

@@

\subsubsection*{Equations of motion}\label{sssec:1D_Model:EquationOfMotion}
\begin{eqnarray}
	\rho A \partial_t^2 w &=& \partial_x V + Q, \label{eq:1D_Model:EquationOfMotion1}\\
	\rho I\partial_t^2 \phi  &=& V + \partial_x M. \label{eq:1D_Model:EquationOfMotion2}
\end{eqnarray}
In \eqref{eq:1D_Model:EquationOfMotion1} and \eqref{eq:1D_Model:EquationOfMotion2} $\rho$ denotes the density, $A$ is the area of a cross section, $I$ is the area moment of inertia, $M$ is the moment, $V$ is the shear force and $Q$ is an external force acting on the beam.

@@

\begin{eqnarray}
	\int_{0}^{1} \partial_{t}^{2} w v  &=& -\int_{0}^{1}(\partial_x w-\phi) v'  + \int_{0}^{1} Q v \nonumber\\
	& & + \left. \partial_x w(\cdot,t) v(\cdot) \right |_0^1 -\left. \phi(\cdot,t) v(\cdot) \right |_0^1, \label{TGT_13}\\
	\int_{0}^{1}\frac{1}{\alpha} \partial_{t}^{2} \phi \psi  &=&  \int_{0}^{1} (\partial_x w - \phi) \psi   - \frac{1}{\beta} \int_{0}^{1}\partial_x \phi \psi' \nonumber \\
	& &+ \frac{1}{\beta}\partial_x \phi(\cdot,t) \psi(\cdot) |_0^1 .\label{TGT_14}
\end{eqnarray}
\subsubsection{Function spaces}\label{sssec:1D_Model:FunctionSpace}
It is convenient to define the following function spaces:
\begin{eqnarray}
	F_0[0,1] & = & \left\{f \in C^1[0,1] \ | \ f(0) = f(1) = 0 \right\} \label{eq:1D_Model:FunctionSpace1}\\
	F_1[0,1] & = & \left\{g \in C^1[0,1] \ | \ g(0) = 0 \right\} \label{eq:1D_Model:FunctionSpace2}
\end{eqnarray}

@@

\textcolor{red}{***********Maak seker L2 is in die appendix!.}\\

>>>>

    \subsection{Equations of motion and constitutive equations}\label{subsec:equationsofmotion+constitutiveequation}
    In this section we introduce the Timoshenko beam theory for the transverse vibration of a uniform beam.
    For a reference of the model~\cite{Inm94}, and~\cite{Fu65} were used.\\

@@

    \subsubsection*{Equations of motion}\label{sssec:1D_Model:EquationOfMotion}
    \begin{align}
        \rho A \partial_t^2 w &=& \partial_x V + Q, \label{eq:1D_Model:EquationOfMotion1}\\
        \rho I\partial_t^2 \phi  &=& V + \partial_x M. \label{eq:1D_Model:EquationOfMotion2}
    \end{align}
    In~\eqref{eq:1D_Model:EquationOfMotion1} and \eqref{eq:1D_Model:EquationOfMotion2} $\rho$ denotes the density, $A$ is the area of a cross section, $I$ is the area moment of inertia, $M$ is the moment, $V$ is the shear force and $Q$ is an external force acting on the beam.

@@

    \begin{eqnarray}
        \int_{0}^{1} \partial_{t}^{2} w v  &=& -\int_{0}^{1}(\partial_x w-\phi) v' + \int_{0}^{1} Q v \nonumber\\
        & & + \left. \partial_x w(\cdot,t) v(\cdot) \right |_0^1 -\left. \phi(\cdot,t) v(\cdot) \right |_0^1, \label{TGT_13}\\
        \int_{0}^{1}\frac{1}{\alpha} \partial_{t}^{2} \phi \psi  &=&  \int_{0}^{1} (\partial_x w - \phi) \psi   - \frac{1}{\beta} \int_{0}^{1}\partial_x \phi \psi' \nonumber \\
        & &+ \frac{1}{\beta}\partial_x \phi(\cdot,t) \psi(\cdot) |_0^1 .\label{TGT_14}
    \end{eqnarray}

    \subsubsection{Function spaces}\label{sssec:1D_Model:FunctionSpace}
    It is convenient to define the following function spaces:
    \begin{eqnarray}
        F_0[0,1] & = & \left\{f \in C^1[0,1] \ | \ f(0) = f(1) = 0 \right\} \label{eq:1D_Model:FunctionSpace1}\\
        F_1[0,1] & = & \left\{g \in C^1[0,1] \ | \ g(0) = 0 \right\} \label{eq:1D_Model:FunctionSpace2}
    \end{eqnarray}

@@

### 2023-06-25 files/Chapter2/generalvibrationmodel.tex
Veranderinge 10-06

The test functions $T[0,1]$ do not form a complete space. Let $V(0,1)$ be the closure of {$F_1(0,1)$} in $H^1(0,1)$. The product space {$V(0,1)\times V(0,1)$} is denoted by $V$ and is complete. It is called the energy space.\\

@@

Let $a$, $b$ and $c$ be bilinear forms where $a$ and $b$ are defined on $V$ and $c$ is defined on $W$. For the models in this dissertation, the bilinear form $b$ is symmetric.

@@

	\item[] \textbf{A2} - There exists a positive constant $C_{W}$ such that $\Vert w\Vert_{X} \leq C_{W}\Vert w\Vert_{W}$ for each $ w\in W$ and $c(\cdot,\ \cdot) =(\cdot,\ \cdot)_{W}$.

	\item[] \textbf{A3} - There exists a positive constant $C_{V}$ such that $\Vert v\Vert_{W} \leq C_{V}\Vert v\Vert_{V}$ for each $v \in V$ and $b(\cdot,\ \cdot) =(\cdot,\ \cdot)_{V}$.

@@

In \cite{VV02} the authors consider a special case of weak damping. Define a space $E_b \subset V$ where \[ E_b = \left\{x \in V \ | \ \textrm{there exists a } y \in W \textrm{ such that } c(y,v) = b(x,v) \textrm{ for all } v \in V\right\}.\] It is proved in \cite{VV02}  that $D(A) = E_b \times V$. Consequently if $u_1 \in V$ and $u_0 \in E_b$ the pair $u_0$, $u_1$ is admissible.

@@

Since $c$ is a bilinear form and by the inequality above, $c$ is a inner-product for $W$. The norm of $W$ can be defined as $||\cdot||_W = \sqrt{c(\cdot,\cdot)}$. Let $C_1 = \min\left\{1, \frac{1}{\alpha}\right\}$ and $C_2 = \max\left\{1, \frac{1}{\alpha}\right\}$. Then
\begin{eqnarray}
	C_1||x||_X \leq ||x||_W \leq C_2||x||_X \label{C}
\end{eqnarray} for all $x \in X$.\\

To prove \textbf{Assumption A3}, some preparation is required. Consider the following proposition for a Poincar\'e-Type inequality for the one-dimensional case.
\newtheorem*{Poincare}{Proposition 1}
\begin{Poincare}
	Suppose that $f \in C^1[a,b]$, and $f(0) = 0$. Then $||f|| \leq (b-a)||f'||$
\end{Poincare}
\begin{proof}
	Let $f \in C^1(0,1)$ such that $f(0) = 0$. By the Fundamental Theorem of Calculus,
	\begin{eqnarray*}
		|f(x)| = \left|\int_0^x f'(s) \ ds\right| \leq \int_0^x |f'(s)| \ ds.
	\end{eqnarray*}
	Since $x$ is arbitrary,
	\begin{eqnarray*}
		||f||_{\sup} \leq \int_0^1 |f'(s)| ds.
	\end{eqnarray*}
	Also for $f,g \in C^1(0,1)$,
	\begin{eqnarray}
		\left| \int_a^b fg \right| & \leq & ||f||\ ||g|| \ \ \ \ \textrm{ by Cauchy-Swartz Inequality.} \label{CSI}
	\end{eqnarray}
	Choose $g = 1$ then since $||f|| \leq ||f||_{\sup} \sqrt{b-a}$, the result follows.
\end{proof}

\newtheorem*{Poincare_2}{Corollary}
\begin{Poincare_2}
	Suppose that $f \in H^1(a,b)$, and $f(0) = 0$. Then $||f|| \leq (b-a)||f'||$.
\end{Poincare_2}
\begin{proof}
	\textcolor{red}{Inlas 41A. Nie seker wat 41A verwys in die veranderinge nie. Kom die ander deel reg maak}
\end{proof}

@@

\begin{proof}
	Let $f \in V$, then $f_1(0) = f_2(0) = 0$. Using the Corollary, and from the definition of the norm $||\cdot||_X$, the following inequality for the norm $||\cdot||_X$ can be obtained.
	\begin{eqnarray*}
		||f||_X^2 & \leq & ||f_1'||^2 + ||f_2'||^2 \\
		& = & ||f_1'||^2 -||f_2||^2 + ||f_2'||^2 +||f_2||^2\\
		& \leq & 2||f_1'-f_2||^2 + 2||f_2'||^2.
	\end{eqnarray*}

@@

	\begin{eqnarray}
		b(f,f) & = &   \int_0^1 (f_1'-f_2)^2 + \frac{1}{\beta}\int_0^1  (f_2')^2 \nonumber\\
		& = & ||f_1' - f_2||^2 + \frac{1}{\beta}||f_2'||^2 \nonumber\\
		& \geq & \min\left\{1 ,\ \frac{1}{\beta} \right\} \left( ||f_1' - f_2||^2 + ||f_2'||^2\right) \label{b}
	\end{eqnarray}

	Now combine the inequalities \eqref{f} and \eqref{b},
	\begin{eqnarray*}
		||f||_W^2 \leq Kb(f,f)
	\end{eqnarray*} for all $f \in V$.
\end{proof}

@@

Therefore $\displaystyle \frac{T''(t)}{T(t)}$ is a constant. Suppose $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$. This can also be written as the following differential equation,
\begin{eqnarray}
	\ddot{T} + \lambda T = 0. \label{eq:1D_Model:ModalAnalysisT}
\end{eqnarray}

@@

Solutions to Problem T-1E is a pair of functions $\langle \tilde{w}, \tilde{\phi} \rangle$ called the eigenfunction with corresponding eigenvalue $\lambda$. Substitution  that $\tilde{w} = \sin(k\pi x)$ and $\tilde{\phi} = \cos(k \pi x)$ are possible solutions of the differential equations and satisfy the boundary conditions.

Due to the periodicity of these solutions, the eigenvalue problem has infinitely many solutions $\langle \tilde{w}_n, \tilde{\phi}_n \rangle$ and corresponding sequence of eigenvalues $\lambda_n$. Another solution is the vector $\langle 0, 1 \rangle$ which also satisfies the eigenvalue problem with boundary conditions. This solution is called the constant solution.

@@

Corresponding to this sequence of eigenvalues, we have a sequence of solutions $(T_n(t))$:
\begin{eqnarray}
	T_k(t)=A_n\cos(\sqrt{\lambda_n}t)+B_n\sin(\sqrt{\lambda_n}t), \label{eq:1D_Model:ModalAnalysisTn}
\end{eqnarray}
(with $A_n$ and $B_n$ arbitrary constants) for the ordinary differential equation \eqref{eq:1D_Model:ModalAnalysisT}.

Combining the solutions of the eigenvalue problem with the solution \eqref{eq:1D_Model:ModalAnalysisTn} yields:
\begin{eqnarray*}
	w_n(x,t) & = & T_n(t)\tilde{w}_n(t), \\
	\phi_n(x,t) & = & T_n(t)\tilde{\phi}_n(x).
\end{eqnarray*} 
Substitution show that these solutions does indeed satisfy the differential equations \ref{eq:1D_Model:ModalPDE1} and \ref{eq:1D_Model:ModalPDE2} with the boundary conditions. These solutions are the modal solutions and they are clearly periodic with natural angular frequencies $\sqrt{\lambda_n}$ (consequently the natural frequencies are $\frac{\sqrt{\lambda_n}}{2 \pi})$.

@@

\subsection{General vibration problem}
This section is a discussion of a general vibration problem. The discussion will refer to the article \cite{CVV18}. In this article, the authors take damping into consideration, which is not covered in this dissertation.

@@

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
\begin{eqnarray}
c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
\end{eqnarray}
with initial conditions $u(0)=u_{0}$ and $u'(0)=u_{1}$.

@@

Therefore $\displaystyle -\frac{T''(t)}{T(t)}$ must be a non-negative constant. Suppose that $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$. The existence of such a $\lambda$ is uncertain at this point. So we consider the following eigenvalue problem.

@@

The article \cite{CVV18} is a convenient reference to show that the eigenvalue problem has a solution if the following assumption is satisfied:

@@

This assumption expands on the assumptions \textbf{A1}-\textbf{A4} already made in Section 2.2.

Building these assumptions, the authors of \cite{CVV18} prove that there exists a eigenvalue $\lambda$ and a corresponding eigenvector $x$ such that the eigenvalue problem has a solution. The authors also prove that the eigenvalues are positive, and the corresponding eigenvectors are orthogonal in $W$ (the bilinear form $c$). In fact it is also orthogonal with respect to the bilinear form $b$, hence
\begin{eqnarray*}
	b(x_i, x_j) = \lambda_i c(x_i, x_j) = 0, \ \ \ \ \textrm{ for each } \ i \neq j.
\end{eqnarray*}

The sequence of eigenvalues ${\lambda_i}$ is an infinite sequence with $\lambda_n \rightarrow \infty$ as $n \rightarrow \infty$. The sequence of normalized eigenvectors ${x_i}$ forms an orthonormal basis in $W$, which means that the sequence is complete (or total) in $W$.

Following this result, the authors explain that any $u \in V$, the eigenvector series for $u$ converges to $u$. These coefficients $a_i$ are generalized Fourier coefficients of $u$ with respect to the basis $x_i$. Therefore, for any $u \in V$,
\begin{eqnarray*}
	u = \sum_{i=1}^{\infty} a_i x_i. = \sum_{i=1}^{\infty} c(u, x_i)x_i.
\end{eqnarray*}

Now that the eigenvalue problem has a solution, the following ordinary differentiable equation can be considered,
\begin{eqnarray*}
	T'' + \lambda T = 0. 
\end{eqnarray*}

Since this differential equation is a simple second order differential equation, T(t) has the following possible solutions:
\begin{flalign}
	T_n(t) &=  A_n \cos(\sqrt{\lambda_n} t) + B_n \sin(\sqrt{\lambda_n} t) \quad \textrm{ if } \lambda_n > 0, \label{lambda_1}\\
	T_n(t) &=  A_n + B_n t \quad \textrm{ if } \lambda_n = 0. \label{lambda_2}
\end{flalign}

@@

\subsubsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we can again refer to the article \cite{CVV18}.

In article, the validity of the series solution is proved using energy norms, where the energy $\mathcal{E}$ a function $u$ given by
\begin{eqnarray}
	\mathcal{E} (t) = \frac{1}{2} b(u(t), u(t)) + \frac{1}{2} c(u'(t), u'(t)). \label{eq:1D_Model:ModalAnalysisEnergy}
\end{eqnarray}

@@

@@

In this dissertation, we are trying to compare different models for use in applications. These comparisons are made in Chapter 4 and 6. Being able to express the solutions of the models as a formal series solution, enable us to compare the different models by only considering the eigenvalues and eigenfunctions of the different models.

>>>>

The test functions $T[0,1]$ do not form a complete space. Let $V(0,1)$ be the closure of {$F_1(0,1)$} in $H^1(0,1)$ (see Section 1.4.4). It follows that $V(0,1)$ is complete and the product space {$V(0,1)\times V(0,1)$} denoted by $V$ is complete. It is called the energy space.\\

@@

Let $a$, $b$ and $c$ be bilinear forms where $a$ and $b$ are defined on $V$ and $c$ is defined on $W$. For the models in this dissertation, the bilinear forms $b$ and $c$ are symmetric. Furthermore, $b(\cdot, \cdot) = (\cdot, \cdot)_V$ and $c(\cdot, \cdot) = (\cdot, \cdot)_W$.

@@

	\item[] \textbf{A2} - There exists a positive constant $C_{W}$ such that $\Vert w\Vert_{X} \leq C_{W}\Vert w\Vert_{W}$ for each $ w\in W$.

	\item[] \textbf{A3} - There exists a positive constant $C_{V}$ such that $\Vert v\Vert_{W} \leq C_{V}\Vert v\Vert_{V}$ for each $v \in V$.

@@

In \cite{VV02} the authors consider a special case of weak damping. Define a space $E_b \subset V$ where \[ E_b = \left\{x \in V \ | \ \textrm{there exists a } y \in W \textrm{ such that } c(y,v) = b(x,v) \textrm{ for all } v \in V\right\}.\] It is proved in \cite{VV02}  that condition (2.2.3) is satisfied if $u_1 \in V$ and $u_0 \in E_b$ (the pair $u_0$, $u_1$ is admissible).

@@

Since $c$ is a bilinear form and by the inequality above, $c$ is an inner-product for $W$. The norm of $W$ is defined as $||\cdot||_W = \sqrt{c(\cdot,\cdot)}$. Let $C_1 = \min\left\{1, \frac{1}{\alpha}\right\}$ and $C_2 = \max\left\{1, \frac{1}{\alpha}\right\}$. Then
\begin{eqnarray}
	C_1||x||_X \leq ||x||_W \leq C_2||x||_X \label{C}
\end{eqnarray} for all $x \in X$.\\

To prove \textbf{Assumption A3}, some preparation is required. Consider the following proposition for a Poincar\'e-Type inequality for the one-dimensional case.
\newtheorem*{Poincare}{Proposition 1}
\begin{Poincare}
	Suppose that $f \in C^1[a,b]$, and $f(a) = 0$. Then\\ $\displaystyle ||f|| \leq (b-a)||f'||$.
\end{Poincare}
\begin{proof}
	Let $f \in C^1(a,b)$ such that $f(a) = 0$. By the Fundamental Theorem of Calculus,
	\begin{eqnarray*}
		|f(x)| = \left|\int_a^x f'(s) \ ds\right| \leq \int_a^x |f'(s)| \ ds.
	\end{eqnarray*}
	Since $x$ is arbitrary,
	\begin{eqnarray*}
		||f||_{\sup} \leq \int_a^b |f'(s)| ds.
	\end{eqnarray*}
	Also for $f,g \in C^1(a,b)$,
	\begin{eqnarray}
		\left| \int_a^b fg \right| & \leq & ||f||\ ||g|| \ \ \ \ \textrm{ by Cauchy-Swartz Inequality.} \label{CSI}
	\end{eqnarray}
	Choose $g = 1$ then since $||f|| \leq ||f||_{\sup} \sqrt{b-a}$, the result follows.
\end{proof}

\newtheorem*{Poincare_2}{Corollary}
\begin{Poincare_2}
	Suppose that $f \in H^1(a,b)$, and $f(a) = 0$. Then $||f|| \leq (b-a)||f'||$.
\end{Poincare_2}
\begin{proof}
	There exists a sequence $(g_n) \subset C^1[a,b]$ such that $||g_n -f|| \rightarrow 0$ and $||g'_n-f'||\rightarrow 0$.

	Therefore $\displaystyle \frac{||f||}{||f'||} = \lim_{n\ \rightarrow \infty} \frac{||g_n||}{||g'_n||} \leq (b-a)$.
\end{proof}

@@

\begin{proof}
	Let $f \in V$, then $f_1(0) = f_2(0) = 0$. Following from the corollary, $||f_1|| \leq ||f_1'||$ and $||f_2|| \leq ||f_2'||$. Therefore,
	\begin{eqnarray*}
		||f_1'|| & = & ||f_1 + f_2 - f_2||,\\
				 & \leq & ||f_1 + f_2|| + ||f_2||.\\
		||f_1'||^2 & \leq & 2||f_1 + f_2||^2 +||f_2'||^2.
	\end{eqnarray*}

	Hence from the definition of the norm $||\cdot||_X$, and again from the corollary,
	\begin{eqnarray*}
		||f||_X^2 & = & ||f_1||^2 + ||f_2||^2,\\
				  & \leq & 2||f_1 + f_2||^2 + 2||f_2'||^2.
	\end{eqnarray*}

@@

	But we also have
	\begin{eqnarray}
		b(f,f) & = &   \int_0^1 (f_1'-f_2)^2 + \frac{1}{\beta}\int_0^1  (f_2')^2, \nonumber\\
		& = & ||f_1' - f_2||^2 + \frac{1}{\beta}||f_2'||^2, \nonumber\\
		& \geq & \min\left\{1 ,\ \frac{1}{\beta} \right\} \left( ||f_1' - f_2||^2 + ||f_2'||^2\right). \label{b}
	\end{eqnarray}

	Now combine the inequalities \eqref{f} and \eqref{b},
	\begin{eqnarray*}
		||f||_W^2 \leq Kb(f,f),
	\end{eqnarray*} for all $f \in V$.
\end{proof}

@@

Therefore $\displaystyle \frac{T''(t)}{T(t)}$ is constant. Suppose $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$. This can also be written as the following differential equation,
\begin{eqnarray}
	\ddot{T} + \lambda T = 0. \label{eq:1D_Model:ModalAnalysisT}
\end{eqnarray}

@@

Solutions to Problem T-1E is a pair of functions $\langle \tilde{w}, \tilde{\phi} \rangle$ called the eigenfunction with corresponding eigenvalue $\lambda$. Substitution of $\tilde{w} = \sin(k\pi x)$ and $\tilde{\phi} = \cos(k \pi x)$ show that they are solutions of the differential equations and satisfy the boundary conditions.

Clearly the eigenvalue problem has infinitely many solutions $\langle \tilde{w}_n, \tilde{\phi}_n \rangle$ and corresponding eigenvalues $\lambda_n$. Another solution is the vector $\langle \tilde{w}, \tilde{\phi} \rangle= \langle 0, 1 \rangle$ which also satisfies the eigenvalue problem with boundary conditions.

@@

Corresponding to this sequence of eigenfunctions, we have a sequence of solutions $(T_n(t))$:
\begin{eqnarray}
	T_n(t)=A_n\cos(\sqrt{\lambda_n}t)+B_n\sin(\sqrt{\lambda_n}t), \label{eq:1D_Model:ModalAnalysisTn}
\end{eqnarray}
(with $A_n$ and $B_n$ arbitrary constants) for the ordinary differential equation \eqref{eq:1D_Model:ModalAnalysisT}.

Combining the solutions of the eigenvalue problem with the solution \eqref{eq:1D_Model:ModalAnalysisTn} yields:
\begin{eqnarray*}
	w_n(x,t) & = & T_n(t)\tilde{w}_n(t), \\
	\phi_n(x,t) & = & T_n(t)\tilde{\phi}_n(x).
\end{eqnarray*} 
Substitution show that these solutions does indeed satisfy the differential equations \eqref{eq:1D_Model:ModalPDE1} and \eqref{eq:1D_Model:ModalPDE2} with the boundary conditions. These solutions are the modal solutions and they are clearly periodic with natural angular frequencies $\sqrt{\lambda_n}$ (consequently the natural frequencies are $\frac{\sqrt{\lambda_n}}{2 \pi})$.

@@

\subsection{General vibration problem}
This section is a discussion of the general vibration problem GVar. The discussion will refer to the article \cite{CVV18}. In this article, the authors take damping into consideration, which is not covered in this dissertation.

@@

\subsubsection*{Problem GVar}\label{sssec:existence:ProblemGVar}
Find a function $u \in C(J,\ X)$ such that $u'$ is continuous at $0$ with respect to $\Vert \cdot \Vert_{W}$, and for each $t \in J$, $u(t) \in V$, $u'(t) \in V$, $u''(t) \in W$, satisfying
\begin{eqnarray}
c(u''(t),v)+b(u(t),v) = 0 \ \ \ \ \textrm{for each} \ v \in V, \label{eq:existence:ProblemGVarHom}
\end{eqnarray}
with initial conditions $u_{0}$ and $u_{1}$.

@@

Therefore $\displaystyle -\frac{T''(t)}{T(t)}$ must be constant. Suppose that $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$. The existence of such a $\lambda$ is uncertain at this point. So we consider the following eigenvalue problem.

@@

The article \cite{CVV18} is a convenient reference to show that the eigenvalue problem has a solution if the following assumption, aditional to \textbf{A1}, \textbf{A2}, \textbf{A3} and \textbf{A4},  is satisfied:

@@

(This assumption expands on the assumptions \textbf{A1}-\textbf{A4} already made in Section 2.2.)

Using these assumptions, the authors of \cite{CVV18} prove that there exists a complete orthonormal sequence of eigenvectors for the eigenvalue problem with a corresponding sequence of eigenvalues. These eigenvalues are positive and the orthogonality is with respect to the bilinear form $c$. In fact it is also orthogonal with respect to the bilinear form $b$,
\begin{eqnarray*}
	b(x_i, x_j) = \lambda_i c(x_i, x_j) = 0, \ \ \ \ \textrm{ for each } \ i \neq j.
\end{eqnarray*}

Also the sequence of normalized eigenvectors ${x_i}$ forms an orthonormal basis in $W$ and sequence of eigenvalues ${\lambda_i}$ is an infinite sequence with $\lambda_n \rightarrow \infty$ as $n \rightarrow \infty$.

Following these results from \cite{CVV18}, for any $u \in V$, $\displaystyle u = \sum_{i=1}^{\infty} a_i x_i$. These coefficients $a_i$ are generalized Fourier coefficients of $u$ with respect to the eigenvectors $x_i$. Therefore, for any $u \in V$,
\begin{eqnarray*}
	u = \sum_{i=1}^{\infty} a_i x_i. = \sum_{i=1}^{\infty} c(u, x_i)x_i.
\end{eqnarray*}

Now that the eigenvalue problem has many solutions, the following ordinary differentiable equation can be considered,
\begin{eqnarray*}
	T_n'' + \lambda T_n = 0. 
\end{eqnarray*}

Since this differential equation is a simple second order differential equation, $T_n(t)$ has the following possible solutions:
\begin{flalign}
	T_n(t) &=  A_n \cos(\sqrt{\lambda_n} t) + B_n \sin(\sqrt{\lambda_n} t) \quad \textrm{ if } \lambda_n > 0, \label{lambda_1}
\end{flalign}

@@

\subsubsection{Validity of series solution}
Consider the following question: When is the formal series solution valid for the vibration problem with initial values $u(0) = u_0$ and $u'(0) = u_1$? To answer this question, we again refer to the article \cite{CVV18}.

In the article, the validity of the series solution is proved using energy norms, where the energy $\mathcal{E}$ a function $u$ given by
\begin{eqnarray}
	\mathcal{E} (t) = \frac{1}{2} b(u(t), u(t)) + \frac{1}{2} c(u'(t), u'(t)). \label{eq:1D_Model:ModalAnalysisEnergy}
\end{eqnarray}

*****************Voorstel*****************************

Then,
\begin{eqnarray*}
	\mathcal{E}'(t) & = & b(u(t), u'(t)) + c(u'(t), u''(t)),\\
					& = & 0 \ \ \ \ \ \ \ \textrm{ following from \eqref{eq:existence:ProblemGVarHom}}. 
\end{eqnarray*}

Therfore $\mathcal{E}(t) = \mathcal{E}(0)$ for all $t>0$. For the case of weak damping, it can be proved that $\mathcal{E}(t) \leq \mathcal{E}(0)$ for all $t>0$ which is the case of \cite{CVV18}.

Denote the partial sum $u^{N}(t) = \sum_{n=1}^{N} T_{n}(t)w_n$. Let
\begin{eqnarray*}
	u_0^{N} = \sum_{n=1}^{N} T_n(0) w_n \ \ \textrm{ and } \ \ u_{1}^{N} =\sum_{n=1}^{N} T'_n(0) w_n.
\end{eqnarray*}

Let $\mathcal{E}$ denote the energy of the error function $u^E_N$. Since $\mathcal{E}(t) = \mathcal{E}(0)$ for all $t>0$ it follows that, \[||u(t) -  u^N(t)||_V^2 + ||u'(t) - (u^N)'(t)||^2_W = ||u_0 - u^N_0||_V^2 + ||u_1 - u^N_1||_W^2. \]

\textcolor{red}{Nie seker wat ek hier moet doen met die Fourier coefficients nie.}

*****************************************************

\textcolor{red}{**************** OU WEERGAWE******************************}

@@

\textcolor{red}{**********************************************}

@@

In this dissertation, the stated objectives were to compare different models for use in applications. These comparisons are made in Chapter 4 and 6. Being able to express the solutions of the models as a formal series solution, enable us to compare the different models by only considering the eigenvalues and eigenfunctions of the different models.

### 2023-06-25 files/Chapter3/dynamicconvergence.tex
Veranderinge 10-06

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, two important convegence results of the Finite Element Method (FEM) is discussed. The first result looks at the convergence of the Galerkin approximation for second order hyperbolic type problems, as presented in \cite{BV13}. The second result, from the textbook \cite{SF97}, examines the convergence of eigenvalues and eigenfunctions for a one-dimensional vibration problem, using the Finite Element Method.

\section{Convergence of the Galerkin approximation}
In the article \cite{BV13}, the authors investigate the convergence of the Galerkin approximation for second order hyperbolic type problems. The article \cite{BV18} extends the work of \cite{BV13} by including general damping and damping at the endpoints. For the models in Chapter 1, \cite{BV13} is sufficient while \cite{BV18} provides more insight and improved notation.

@@

\subsection{Galerkin approximation of Problem GVar}
Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in the Hilbert space $H$.

>>>>

\documentclass[../../main.tex]{subfiles}
\insubfile{\AtEndDocument{\bibliography{../../references.bib}}}

\begin{document}
\section{Introduction}
In this chapter, two important convergence results of the Finite Element Method
(FEM) is discussed. The first result looks at the convergence of the Galerkin
approximation for second order hyperbolic type problems, as presented
in~\cite{BV13}. The second result, from the textbook~\cite{SF97}, examines the
convergence of eigenvalues and eigenfunctions for a one-dimensional vibration
problem, using the Finite Element Method.

\section{Convergence of the Galerkin approximation}\label{sec:convergence-of-the-galerkin-approximation}
In the article \cite{BV13}, the authors investigate the convergence of the
Galerkin approximation for second order hyperbolic type problems. The article
\cite{BV18} extends the work of \cite{BV13} by including general damping and
damping at the endpoints. For the models in Chapter 1, \cite{BV13} is
sufficient while \cite{BV18} provides more insight and improved notation.

@@

\subsection{Galerkin approximation of Problem GVar}\label{subsec:galerkin-approximation-of-problem-gvar}
Recall from Chapter 2 that $V$ is the closure of the test function space $T$ in
the Hilbert space $H$.
